---
title: "EDS241: Assignment 4"
author: "Jake Eisaguirre"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

librarian::shelf(here, tidyverse, janitor, estimatr, kableExtra, broom, car, AER)
```


##### Clean Data

\noindent The following code loads and cleans the data.
```{r}
data <- read_csv(here("EU_sardines.csv")) %>% 
  clean_names()
```

##### (a) Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.

\noindent The following code estimates the bivariate regression and price elasticity demand for sardines
```{r}

mod1 <- lm_robust(log(volume_sold_kg) ~ log(price_euro_kg), data = data)

tidy(mod1) %>% 
  kable()

```



##### (b) Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic. 

```{r}

fsl <- lm(log(price_euro_kg) ~ wind_m_s, data = data)
summary(fsl)

linearHypothesis(fsl, c("wind_m_s = 0"), white.adjust = "hc2")

```



##### (c) Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines? 

```{r}

tsls <- ivreg(log(volume_sold_kg) ~ log(price_euro_kg) | wind_m_s, data = data)
summary(tsls)



```



##### (d) Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.

```{r}

tsls3 <- ivreg(log(volume_sold_kg) ~ log(price_euro_kg) | wind_m_s + as.factor(country) + 
                 as.factor(year) + as.factor(month), data = data)

summary(tsls3)

```

